---
title: "Getting Data into PhyloSeq Object "
output: html_notebook
---
```{r}
#if(!requireNamespace("BiocManager")){
#  install.packages("BiocManager")
#}
#BiocManager::install("phyloseq")
#install.packages("stringr") 
#install.packages("remotes")
#remotes::install_github("jbisanz/qiime2R")              
```


```{r}
library(phyloseq) 
#packageVersion("phyloseq")
#citation("phyloseq")
library(vegan) 
library(tidyverse) 
library(devtools)
library(qiime2R) 
#install.packages("stringr") 
library("stringr")                
```

```{r}
# Importing ASVs abundance file
ASVs <- read_qza("data/2022mar15/table.qza")
ASVs$data

# Importing metadata
metadata <- read.table ("data/2022mar15/mapfile.txt", sep='\t', header=T, row.names=1, comment="")
metadata <- metadata[-1,] # remove the second line that specifies the data type
metadata

metadata$LongitudeStart #changing from character to numeric before creating phyloseq Object
is.factor(metadata$LongitudeStart)
factor <- as.factor(metadata$LongitudeStart)
is.factor(factor)
factor
metadata$LongitudeStart <- factor
is.factor(metadata$LongitudeStart)

# Importing tree
tree <- read_qza("data/2022mar15/rooted-tree.qza")
#tree$data

# Importing taxonomy
taxonomy <- read_qza("data/2022mar15/taxonomy.qza")
#taxonomy$data

tax_table <- do.call(rbind, strsplit(as.character(taxonomy$data$Taxon), ";"))
colnames(tax_table) <- c("Kingdom","Phylum","Class","Order","Family","Genus","Species")
rownames(tax_table) <- taxonomy$data$Feature.ID
#tax_table
```

```{r}
# Creating phyloseq object
ps <- phyloseq(
  otu_table(ASVs$data, taxa_are_rows = TRUE),
  phy_tree(tree$data),
  tax_table(tax_table),
  sample_data(metadata)
)

ntaxa(ps)
nsamples(ps)
sample_names(ps)[1:5]  
rank_names(ps)  
sample_variables(ps)  
otu_table(ps)[1:5, 1:5]  
tax_table(ps)[1:5, 1:4]
```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

